Extracting sunbeams from cucumbers: Reproducibility as a necessity for the future of archaeological science

### 2025-01-13 paper.qmd
results: variation among journals

boxlplot_rel_title_length <- 
items_df_title %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(1,
             relative_title_length)) +
  geom_boxplot(
               size = 1)  +
  # median for physics, from Fanelli & Glänzel Fig 2
  geom_hline(yintercept = 0.8,
             colour = "grey70") +
  geom_text_repel(data = tibble(x = 1, 
                                y = 0.8,
                                label = "p"),
                  aes(x, y, label = label),
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
  # median for soc sci, from Fanelli & Glänzel Fig 2
  geom_hline(yintercept = 0.0,
             colour = "grey70") +
  geom_text_repel(data = tibble(x = 1, 
                                y = 0.0,
                                label = "s"),
                  aes(x, y, label = label),
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
    # median for humanities, from Fanelli & Glänzel Fig 2
  geom_hline(yintercept = -0.2,
             colour = "grey70") +
  geom_text_repel(data = tibble(x = 1, 
                                y = -0.2,
                                label = "h"),
                  aes(x, y, label = label),
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +

@@

```{r}
#| label: fig-compare-other-fields
#| fig-cap: Distributions of article characteristics hypothesised to reflect the level of consensus. The boxplot shows the distribution of values of archaeology articles. The thick black line in the middle of the boxplot is the median value, the box represents the inter-quartile range (the range between the 25th and 75th percentiles, where 50% of the data are located), and individual points represent outliers. The thin grey horizontal lines in each boxplot indicate the median values computed by Fanelli and Glanzel (2013), where p = physics, s = social sciences, h = humanities.

@@

over_time <-  
  items_df %>%  
  left_join(items_df_title) %>%  
  left_join(shannon_per_item) %>%  
  mutate(log_authors_n = log(authors_n), 
         log_pages_n = log(pages_n), 
         journal_wrp = str_wrap(journal, 30)) %>%  
  select(year, 
         log_authors_n, 
         log_pages_n, 
         prices_index,  
         shannon, 
         sqrt_refs_n,  
         relative_title_length)

@@

```{r}
#| label: fig-change-over-time
#| fig-cap: Distribution of article characteristics for archaeology articles over time. D ata points represent individual articles. The Ordinary Least Squares method was used to fit the straight lines summarising the relationships between the variables and the time series.

@@

>>>>

boxlplot_rel_title_length <- 
items_df_title %>% 
  filter(!is.na(year)) %>% 
  ggplot(aes(1,
             relative_title_length)) +
  geom_boxplot(
               size = 1)  +
  # median for physics, from Fanelli & Glänzel Fig 2
  geom_hline(yintercept = 0.8,
             colour = "grey70") +
  geom_text_repel(data = tibble(x = 1, 
                                y = 0.8,
                                label = "p"),
                  aes(x, y, label = label),
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0,
                  nudge_x = 0.05) +
  # median for soc sci, from Fanelli & Glänzel Fig 2
  geom_hline(yintercept = 0.0,
             colour = "grey70") +
  geom_text_repel(data = tibble(x = 1, 
                                y = 0.0,
                                label = "s"),
                  aes(x, y, label = label),
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +
    # median for humanities, from Fanelli & Glänzel Fig 2
  geom_hline(yintercept = -0.2,
             colour = "grey70") +
  geom_text_repel(data = tibble(x = 1, 
                                y = -0.2,
                                label = "h"),
                  aes(x, y, label = label),
                  bg.colour = "white", 
                  bg.r = .2, 
                  force = 0) +

@@

```{r}
#| label: fig-compare-other-fields
#| fig-cap: Distributions of article characteristics hypothesised to reflect the level of consensus. The boxplot shows the distribution of values of archaeology articles. The thick black line in the middle of the boxplot is the median value, the box represents the inter-quartile range (the range between the 25th and 75th percentiles, where 50% of the data are located), and individual points represent outliers. The thin grey horizontal lines in each boxplot indicate the median values computed by Fanelli and Glanzel (2013), where p = physics, s = social sciences, h = humanities. 
#| fig-height: 10

@@

over_time <-  
  items_df %>%  
  left_join(items_df_title) %>%  
  left_join(shannon_per_item) %>%  
  mutate(log_authors_n = log(authors_n), 
         log_pages_n = log(pages_n), 
         journal_wrp = str_wrap(journal, 30)) %>%  
  select(year, 
         log_authors_n, 
         log_pages_n, 
         prices_index,  
         shannon, 
         relative_title_length)

@@

```{r}
#| label: fig-change-over-time
#| fig-cap: Distribution of article characteristics for archaeology articles over time. D ata points represent individual articles. The Ordinary Least Squares method was used to fit the straight lines summarising the relationships between the variables and the time series. 
#| fig-height: 10

@@

```{r}
library(ggridges)

journal_title_size <- 4

plot_journals_authors <- 
items_df %>% 
  mutate(journal = str_wrap(journal, 20)) %>% 
  ggplot(aes(y = reorder(journal,
                     authors_n), 
             x = log(authors_n),
             fill =   after_stat(x),
             height = after_stat(density))) +
  geom_density_ridges_gradient(stat = "density",
                               colour = "white") +
  scale_fill_viridis_c() +
  guides(fill = 'none') +
  theme_minimal() +
  theme(axis.text.y = element_text(size = journal_title_size)) +
  ylab("") + 
  xlab("Number of authors (ln)")

plot_journals_article_length <- 
items_df %>% 
  mutate(journal = str_wrap(journal, 20)) %>% 
  ggplot(aes(y = reorder(journal,
                     -pages_n), 
             x = log(pages_n),
             fill =   after_stat(x),
             height = after_stat(density))) +
  geom_density_ridges_gradient(stat = "density",
                               colour = "white") +
  scale_fill_viridis_c() +
  guides(fill = 'none') +
  theme_minimal() +
  theme(axis.text.y = element_text(size = journal_title_size)) +
  xlab("Number of pages (ln)") +
  ylab("")

plot_journals_title_length <- 
items_df_title %>% 
  mutate(journal = str_wrap(journal, 20)) %>% 
  ggplot(aes(y = reorder(journal,
                     relative_title_length), 
             x = relative_title_length,
             fill =   after_stat(x),
             height = after_stat(density))) +
  geom_density_ridges_gradient(stat = "density",
                               colour = "white") +
  scale_fill_viridis_c() +
  guides(fill = 'none') +
  theme_minimal() +
  theme(axis.text.y = element_text(size = journal_title_size)) +
  ylab("") +
  xlab("Relative title length (ln)")

plot_journals_ref_recency <- 
items_df %>% 
  mutate(journal = str_wrap(journal, 20)) %>% 
  group_by(journal) %>% 
  ggplot(aes(y = reorder(journal,
                     prices_index), 
             x = prices_index,            
             fill =   after_stat(x),
             height = after_stat(density))) +
  geom_density_ridges_gradient(stat = "density",
                               colour = "white") +
  scale_fill_viridis_c() +
  guides(fill = 'none') +
  theme_minimal() +
  theme(axis.text.y = element_text(size = journal_title_size)) +
  ylab("") +
  xlab("Recency of references")

plot_journals_ref_diversity <- 
shannon_per_item %>%  
  mutate(journal = str_wrap(journal, 20)) %>% 
  group_by(journal) %>% 
  ggplot(aes(y = reorder(journal,
                     -shannon), 
             x = shannon,
             fill =   after_stat(x),
             height = after_stat(density))) +
  geom_density_ridges_gradient(stat = "density",
                               colour = "white") +
  scale_fill_viridis_c() +
  guides(fill = 'none') +
  theme_minimal() +
  theme(axis.text.y = element_text(size = journal_title_size)) +
  ylab("") +
  xlab("Diversity of references")
```

```{r}
#| label: fig-variation-by-journal
#| fig-cap: Variation by journal
#| fig-height: 10

plot_grid(plot_journals_authors, 
          plot_journals_article_length,
          plot_journals_title_length,
          plot_journals_ref_recency, 
          plot_journals_ref_diversity,
          nrow = 2)
```

@fig-variation-by-journal shows
